### 2023-09-25 metadata/parametadata_control.Rmd
Adding static standards sheet

---
title: "Metadata collation"
author: "William Kumler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
library(RaMS)
```

## Read in water column .gof file

```{r download raw data, eval=!file.exists("metadata/raw_data/paragon1.gof")}
sum_url <- "http://scope.soest.hawaii.edu/data/scope2021/paragon1.sum"
download.file(sum_url, destfile = "metadata/raw_data/paragon1.sum")

gof_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/paragon1.gof"
download.file(gof_url, destfile = "metadata/raw_data/paragon1.gof")

seaflow_url <- "http://scope.soest.hawaii.edu/data/scope2021/SeaFlow_dataset_v1.4.xlsx"
download.file(seaflow_url, destfile = "metadata/raw_data/SeaFlow_dataset_v1.4.xlsx", mode = "wb")

# register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
# gm <- get_map(location = c(-156.5, 21.8), zoom = 8, maptype = "hybrid")
# saveRDS(object = gm, file = "metadata/raw_data/ggmap_of_aloha.rds")

magick::image_read("http://scope.soest.hawaii.edu/data/scope2021/Paragon1CTDchlp.pdf", density = 300) %>%
  magick::image_trim() %>%
  magick::image_write("metadata/raw_data/chl_pdf_url.png", format = "png")
```

![](metadata/raw_data/chl_pdf_url.png)

```{r From SCOPE FTP}
para_cast_times <- "metadata/raw_data/paragon1.sum" %>%
  read_table(skip = 3, col_names = FALSE, col_types = cols(.default = col_character())) %>%
  mutate(mdyhms=paste(X3, X4, X5, X6)) %>%
  mutate(time=as.POSIXct(strptime(mdyhms, "%b %d  %Y %X", tz="UTC"))) %>%
  select(station=X1, cast=X2, time) %>%
  mutate(station=as.numeric(station), cast=as.numeric(cast))

gof_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")
qual_names <- setdiff(gof_names, c(head(gof_names, 6), tail(gof_names, 4)))

# From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
# Quality Indicators:
# Flag  Meaning
# 1    not quality controled
# 2    good data
# 3    suspect (i.e.  questionable) data
# 4    bad data
# 5    missing data
# 9    variable not measured during this cast
raw_gof <- read_table("metadata/raw_data/paragon1.gof", skip = 4, col_names = gof_names,
                      col_types = cols(.default = col_double()))
qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
  unite("value", sep = "") %>%
  separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
  mutate(across(everything(), as.numeric))
raw_gof[,qual_names][qual_df!=2] <- NA
para_core <- raw_gof %>% 
  left_join(para_cast_times) %>%
  select(station:rosette, time, lat:oxy_um, chl_ug:phaeo_ug) %>%
  mutate(lon=-lon)
```

```{r visualize}
my_casts <- c(4, 5, 9, 11, 13, 14, 16, 18, 25, 27, 31, 32, 34, 36, 39, 41)

gm <- readRDS("metadata/raw_data/ggmap_of_aloha.rds")
stations_toplot <- para_core %>%
  distinct(station, cast, lat, lon) %>%
  mutate(station_cast=paste0("S", station, "C", cast)) %>%
  mutate(plotcol=ifelse(cast%in%my_casts, "red", "black"))
satmap <- ggmap(gm) +
  geom_path(aes(x=lon, y=lat), data=stations_toplot, color="grey80") +
  geom_point(aes(x=lon, y=lat), data = stations_toplot, color="white") +
  scale_color_identity()
statmap <- ggplot(stations_toplot, aes(x=lon, y=lat)) +
  geom_path(color="black") +
  geom_text(aes(x=lon, y=lat, color=plotcol, label=station_cast)) +
  scale_color_identity() +
  theme_bw() +
  coord_fixed()
cowplot::plot_grid(satmap, statmap, nrow=1)

para_core %>%
  mutate(station_cast=paste0("S", station, "C", cast)) %>%
  mutate(plotcol=ifelse(cast%in%my_casts, "red", "black")) %>%
  count(plotcol, station_cast, time) %>%
  ggplot(aes(x=time, y=n, fill=plotcol, label=station_cast)) +
  geom_col() +
  scale_fill_identity() +
  scale_x_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d") +
  theme_bw()

para_core %>%
  pivot_longer(temp:phaeo_ug) %>%
  filter(pressure<260) %>%
  drop_na() %>%
  ggplot(aes(x=value, y=-pressure, group=cast, color=time)) +
  geom_point() +
  facet_wrap(~name, nrow = 2, scales="free_x") +
  scale_color_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d")

seaflowdata <- readxl::read_excel("metadata/raw_data/SeaFlow_dataset_v1.4.xlsx")
seaflowdata %>%
  mutate(time=str_remove(time, "\\+.*")) %>%
  mutate(time=strptime(time, format = "%Y-%m-%dT%H:%M:%S")) %>%
  mutate(time=as.POSIXct(time)) %>%
  select(time, starts_with("biomass")) %>%
  pivot_longer(-time) %>%
  filter(!(name=="biomass_synecho" & value>0.2)) %>%
  ggplot() +
  geom_point(aes(x=time, y=value)) +
  facet_wrap(~name, ncol=1, scales="free_y") +
  scale_x_datetime(timezone = "HST", date_breaks = "1 day", date_labels="%b %d")
```

## Create file data from MS samples

```{r}
samp_filedata <- "mzMLs" %>%
  list.files(pattern = "Smp.*mzML", recursive = TRUE, full.names = TRUE) %>% 
  data.frame(filepath=.) %>%
  mutate(filename=basename(filepath)) %>%
  mutate(shortname=str_remove(filename, "-wIS|-noIS-neg|-noIS-pos")) %>%
  mutate(shortname=str_remove(shortname, ".*_Smp_")) %>%
  mutate(samp_type="Smp") %>%
  mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
  mutate(polarity=str_extract(filepath, "neg|pos")) %>%
  mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
  mutate(depth=str_extract(shortname, "Deep|Surf")) %>%
  mutate(startime=str_extract(shortname, "Eve|Morn")) %>%
  mutate(timepoint=str_extract(shortname, "T\\d+")) %>%
  mutate(tripl=str_extract(shortname, "[A-C](?=\\.mzML$)"))

station_samp_map <- tribble(
  ~amendment, ~depth, ~startime, ~cast,
  "Nit", "Surf", "Morn", 4,
  "Nit", "Surf", "Eve", 5,
  "Nit", "Deep", "Morn", 9,
  "Nit", "Deep", "Eve", 11,
  "Arg", "Surf", "Morn", 13,
  "Arg", "Surf", "Eve", 14,
  "Arg", "Deep", "Morn", 16,
  "Arg", "Deep", "Eve", 18,
  "GMP", "Surf", "Morn", 25,
  "GMP", "Surf", "Eve", 27,
  "GMP", "Deep", "Morn", 31,
  "GMP", "Deep", "Eve", 32,
  "Amm", "Surf", "Morn", 34,
  "Amm", "Surf", "Eve", 36,
  "Amm", "Deep", "Morn", 39,
  "Amm", "Deep", "Eve", 41
) %>%
  mutate(station=2)

other_filedata <- "mzMLs" %>%
  list.files(recursive = TRUE, full.names = TRUE, pattern = ".mzML") %>%
  str_subset("Smp.*mzML", negate = TRUE) %>%
  str_subset("MSMS", negate = TRUE) %>%
  str_subset("TMAO", negate = TRUE) %>%
  data.frame(filepath=.) %>% 
  mutate(filename=basename(filepath)) %>%
  mutate(shortname=str_remove(filename, "-wIS|-noIS-neg|-noIS-pos")) %>%
  mutate(shortname=str_remove(shortname, ".*_Poo_|.*_Std_(4uMStds)?")) %>%
  mutate(samp_type=str_extract(filename, "Poo|Std")) %>%
  mutate(IS_type=str_extract(filepath, "wIS|noIS")) %>%
  mutate(polarity=str_extract(filepath, "neg|pos")) %>%
  mutate(amendment=str_extract(shortname, "Nit|Amm|Arg|GMP")) %>%
  fill(IS_type, amendment, .direction = "down")

all_filedata <- samp_filedata %>%
  bind_rows(other_filedata) %>% 
  left_join(station_samp_map)
```

## Get file timestamps using RaMS

```{r}
if(!file.exists("metadata/raw_data/mzml_timestamps.csv")){
  mzml_timestamps <- grabMSdata(all_filedata$filepath, grab_what = "metadata") %>%
    pluck("metadata") %>%
    select(filename, timestamp)
  write_csv(mzml_timestamps, "metadata/raw_data/mzml_timestamps.csv")
}
mzml_timestamps <- read_csv("metadata/raw_data/mzml_timestamps.csv") %>%
  distinct()
parametadata <- all_filedata %>%
  left_join(mzml_timestamps, by="filename")
```

## Write out clean parametadata file

```{r write out}
write_csv(parametadata, file = "metadata/parametadata.csv")
```

## Create a clean and static list of standards

```{r stans}
base_url <- "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/"
mix_stans <- base_url %>%
  paste0("1ddd97186fa9df1646941aec903dc64a65ef6cfe/Ingalls_Lab_Standards.csv") %>%
  read_csv() %>%
  select(compound_name=Compound_Name, mix=HILIC_Mix) %>%
  drop_na() %>%
  distinct()
custom_stans <- base_url %>%
  paste0("master/data_extra/custom_stans.csv") %>%
  read_csv() %>%
  select(compound_name=Compound_Name, mz, rt=RT_minute, z,
         conc_um=Concentration_uM, compound_type=Compound_Type,
         formula=Empirical_Formula, date_added=Date_Added, mix=HILIC_Mix) %>%
  mutate(column="HILIC") %>%
  mutate(date_added=230101) %>%
  mutate(compound_name=str_remove(compound_name, "\\?$"))

all_stans <- paste0(base_url, "master/Ingalls_Lab_Standards.csv") %>%
  read_csv() %>%
  select(compound_name=Compound_Name, mz, rt=RT_minute, z, column=Column, 
         conc_um=Concentration_uM, compound_type=Compound_Type,
         formula=Empirical_Formula, date_added=Date_Added) %>%
  filter(column=="HILIC") %>%
  left_join(mix_stans, by="compound_name") %>%
  bind_rows(custom_stans) %>%
  mutate(date_added=ifelse(is.na(date_added), 230101, date_added)) %>%
  mutate(compound_name=ifelse(date_added<210813, compound_name,
                              paste0(compound_name, "?"))) %>%
  mutate(polarity=ifelse(z<0, "neg", "pos")) %>%
  arrange(mz, rt)
write_csv(all_stans, "metadata/all_stans.csv")
```