### 2025-05-05 targeted/paratargeted_control.Rmd
Adding targeted analysis

---
title: "PARAGON targeted analysis"
author: "William Kumler"
date: "2024-05-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RaMS)
library(arrow)
library(ggh4x)
options(pillar.sigfig=7)

parametadata <- read_csv("metadata/parametadata.csv") %>%
  mutate(depth=fct_inorder(depth)) %>%
  mutate(startime=fct_inorder(startime)) %>%
  mutate(timepoint=fct_inorder(timepoint))
all_stans <- read_csv("metadata/all_stans.csv") %>%
  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
  mutate(n_C=case_when(
    !str_detect(formula, "C")~0,
    str_detect(formula, "C") & is.na(n_C)~1,
    TRUE~n_C
  )) %>%
  mutate(n_N=as.numeric(str_extract(formula, "(?<=N)\\d+"))) %>%
  mutate(n_N=case_when(
    !str_detect(formula, "N")~0,
    str_detect(formula, "N") & is.na(n_N)~1,
    TRUE~n_N
  ))
iso_stans <- all_stans %>%
  select(compound_name, mz, n_C, n_N, polarity) %>%
  pmap(function(...){
    row_data <- list(...)
    expand_grid(n_C=0:row_data$n_C, n_N=0:row_data$n_N, 
                amendment=c("Amm", "Nit", "Arg", "GMP")) %>%
      mutate(compound_name=row_data$compound_name, mz=row_data$mz,
             polarity=row_data$polarity)
  }) %>%
  bind_rows() %>%
  filter(!(amendment%in%c("Amm", "Nit") & n_C>0)) %>%
  mutate(mz=mz+1.003355*n_C+0.997035*n_N) %>%
  mutate(iso_name = paste0(compound_name, ", 13C", n_C, ", 15N", n_N))

splividis <- function(n, ...){
  vir <- viridis::viridis(n, ...)
  if(n%%2==0){
    vir[as.numeric(matrix(1:n, ncol=2, byrow = TRUE))]
  } else {
    index_vec <- suppressWarnings(as.numeric(matrix(1:n, ncol=2, byrow = TRUE)))
    vir[head(index_vec, -1)]
  }
}
```

## Integrations for isotopes
### Make graphs for peak boundaries

```{r extract EICs and convert to PDF using 0-20 RT}
all_stans <- read_csv("metadata/all_stans.csv") %>%
  mutate(polarity=factor(polarity, levels=c("pos", "neg"))) %>%
  arrange(polarity) %>%
  filter(compound_type!="Internal Standard")

all_stans %>%
  select(polarity, compound_name, mz, rt, date_added, mix) %>%
  expand_grid(amendment=c("Amm", "Nit", "GMP", "Arg"), .) %>%
  mutate(rtstart=0, rtend=20, notes=NA) %>%
  write_csv("targeted/peak_bounds.csv")

pdf("targeted/feat_bounds.pdf", onefile = TRUE, width = 7, height = 4)
all_stans %>%
  expand_grid(amendment=c("Amm", "Nit", "GMP", "Arg"), .) %>%
  # head() %>%
  pwalk(function(...){
    row_data <- list(...)
    eic_i <- open_dataset("tmzMLs/pqds") %>%
      filter(polarity==row_data$polarity) %>%
      filter(samp_type%in%c("Poo", "Std")) %>%
      filter(amendment==row_data$amendment) %>%
      filter(mz%between%pmppm(row_data$mz, 10)) %>%
      filter(rt%between%c(0.5, 20)) %>%
      collect()
    gp <- eic_i %>%
      mutate(mixtype=str_extract(filename, "Poo|Mix1|Mix2")) %>%
      filter(!str_detect(filename, "H2OInMatrix")) %>%
      left_join(parametadata, by = join_by(filename, samp_type, amendment, polarity)) %>%
      mutate(mixtype=factor(mixtype, levels=c("Poo", "Mix1", "Mix2"))) %>%
      ggplot() +
      geom_vline(xintercept = 1:19, color="grey70") +
      geom_vline(xintercept = row_data$rt, color="red") +
      geom_line(aes(x=rt, y=int, group=filename, color=mixtype), linewidth=0.8) +
      facet_wrap(~mixtype, ncol=1, scales="free") +
      ggtitle(paste0(row_data$amendment, " ", row_data$compound_name, 
                     "  (", row_data$mix, ")")) +
      theme_bw() +
      theme(legend.position = "none") +
      scale_x_continuous(breaks = 0:20, limits = c(0, 20)) +
      scale_color_manual(breaks=c("Mix1", "Mix2", "Poo"), 
                         values = c("purple", "mediumorchid1", "blue"))
    print(gp)
  }, .progress = TRUE)
dev.off()
```

```{r check integration consistency between runs}
bound_stans <- readxl::read_excel("targeted/peak_bounds.xlsx")

bound_stans %>%
  filter(rtstart>0) %>%
  ggplot() +
  geom_segment(aes(x=rtstart, xend=rtend, y=mz, yend=mz, color=polarity)) +
  facet_wrap(~amendment)

library(GGally)
bound_stans %>%
  filter(polarity=="pos") %>%
  select(compound_name, amendment, rtstart) %>%
  pivot_wider(names_from = amendment, values_from = rtstart) %>%
  column_to_rownames("compound_name") %>%
  ggpairs(aes(alpha=0.5), diag = list(continuous = "blankDiag"))

bound_stans %>%
  filter(polarity=="pos") %>%
  select(compound_name, amendment, rtstart) %>%
  pivot_wider(names_from = amendment, values_from = rtstart) %>%
  filter(Amm>0|GMP>0) %>%
  ggplot(aes(x=Amm, y=GMP, label=compound_name)) +
  geom_abline(slope=1) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point() +
  geom_text() +
  scale_x_continuous(expand = expansion(0.2))
```

```{r check integration bounds by plotting chroms}
bound_stans <- readxl::read_excel("targeted/peak_bounds.xlsx")

eic_data <- bound_stans %>%
  filter(rtstart>0) %>%
  # filter(compound_name%in%head(unique(compound_name))) %>%
  nest(row_data=c(amendment, rtstart, rtend, in_pooled, notes)) %>%
  mutate(eic=pbapply::pbmapply(function(mz_i, polarity_i){
    open_dataset("tmzMLs/pqds") %>%
      filter(polarity==polarity_i) %>%
      filter(mz%between%pmppm(mz_i, 10)) %>%
      filter(rt%between%c(0.5, 20)) %>%
      collect()
    }, .$mz, .$polarity, SIMPLIFY = FALSE))
saveRDS(eic_data, "targeted/eic_data.rds")

eic_data <- readRDS("targeted/eic_data.rds")
pdf("targeted/samp_int_check.pdf", onefile = TRUE, width = 7, height = 4)
for(i in 1:nrow(eic_data)){
  gp <- left_join(eic_data$row_data[[i]], eic_data$eic[[i]], by=join_by(
    amendment, dplyr::between(y$rt, x$rtstart, x$rtend))
    ) %>%
    filter(samp_type%in%c("Smp","Poo")) %>%
    qplotMS1data() +
    ggtitle(eic_data$compound_name[[i]]) +
    facet_wrap(~amendment, scales="free_y")
  print(gp)
}
dev.off()
```

### Integrate areas

```{r get feature eics for integration}
cmpd_bounds <- readxl::read_xlsx("targeted/peak_bounds.xlsx") %>%
  filter(rtstart>0) %>%
  filter(is.na(notes) | notes!="Remove") %>%
  mutate(mz=ifelse(compound_name=="Homarine", 138.0541, mz)) %>%
  select(amendment, polarity, compound_name, rtstart, rtend)

# Update iso_stans to reflect the changes made during the Excel annotation
updated_iso_stans <- iso_stans %>%
  filter(!compound_name=="beta-Glutamic acid") %>%
  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
  filter(!compound_name=="L-Isoleucine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
  filter(!compound_name=="L-Homoserine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
  filter(!compound_name=="Allopurinol") %>%
  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol")) %>%
  distinct(amendment, compound_name, iso_name, polarity, mz, n_C, n_N) %>%
  filter(compound_name%in%unique(cmpd_bounds$compound_name))
new_pory_entry <- updated_iso_stans %>% 
  filter(compound_name=="Porphyra-334?") %>% 
  mutate(compound_name="Also Porphyra-334?") %>%
  mutate(iso_name=str_replace(iso_name, "Porphyra-334\\?", "Also Porphyra-334\\?"))
new_paly_entry <- updated_iso_stans %>% 
  filter(compound_name=="Palythene/Usujirene?") %>% 
  mutate(compound_name="Another Palythene/Usujirene?") %>%
  mutate(iso_name=str_replace(iso_name, "Palythene/Usujirene\\?", "Another Palythene/Usujirene\\?"))
updated_iso_stans <- updated_iso_stans %>% 
  bind_rows(new_pory_entry) %>% 
  bind_rows(new_paly_entry)

final_bounds <- cmpd_bounds %>%
  left_join(updated_iso_stans)

cmpd_eics <- final_bounds %>%
  nest(amend_specs=-c(compound_name, polarity)) %>%
  # head() %>%
  # slice(1)  -> row_data
  # filter(compound_name=="Theanine?") -> row_data
  pmap(function(...){
    row_data <- list(...)
    # print(row_data$compound_name)
    amend_specs <- row_data$amend_specs %>%
      rowwise() %>%
      mutate(mzmin=pmppm(mz, 10)[1]) %>%
      mutate(mzmax=pmppm(mz, 10)[2]) %>%
      ungroup()
    mzmin <- min(amend_specs$mz)-0.5
    mzmax <- max(amend_specs$mz)+0.5
    rtmin <- min(amend_specs$rtstart)
    rtmax <- max(amend_specs$rtend)
    all_eic <- open_dataset("tmzMLs/pqds") %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(mzmin, mzmax)) %>%
      filter(rt%between%c(rtmin, rtmax)) %>%
      collect()
    cmpd_eic <- left_join(amend_specs, all_eic, by=join_by(
      amendment, dplyr::between(y$rt, x$rtstart, x$rtend), 
      dplyr::between(y$mz, x$mzmin, x$mzmax)
    )) %>%
      distinct() %>%
      group_by(amendment, filename, iso_name, rt) %>%
      # nest() %>% ungroup() %>% slice(1) %>% unnest(data)
      slice_max(int) %>%
      ungroup() %>%
      mutate(compound_name=row_data$compound_name) %>%
      select(amendment, compound_name, iso_name, rt, mz=mz.y, int, filename, polarity)
    cmpd_eic
  }, .progress = TRUE) %>%
  bind_rows() %>%
  ungroup()
saveRDS(cmpd_eics, "targeted/cmpd_eics.rds")
```

```{r single compound demo check}
cmpd_eics <- readRDS("targeted/cmpd_eics.rds")
cmpd_eics %>%
  ungroup() %>%
  filter(compound_name=="Theanine?") %>%
  left_join(iso_stans %>% select(-mz)) %>%
  # filter(filename=="210916_Poo_TruePooAmm-noIS-pos-Full_5.mzML") %>%
  filter(n_C==0 & n_N<=1) %>%
  left_join(parametadata) %>%
  mutate(iso_name=str_remove(iso_name, ".*?, ")) %>%
  qplotMS1data(color_col = "timepoint") + 
  facet_grid(iso_name~amendment, scales = "free")
```

```{r integrate areas}
cmpd_eics <- readRDS("targeted/cmpd_eics.rds")
raw_areas <- cmpd_eics %>%
  left_join(updated_iso_stans %>% select(-mz)) %>%
  split(.$compound_name) %>%
  pbapply::pbsapply(function(cmpd_eic_i){
    unlabeled_eic <- cmpd_eic_i %>%
      group_by(filename) %>%
      filter(n_C==0 & n_N==0) %>%
      mutate(unlab_n=n()) %>%
      select(filename, rt, unlab_int=int, unlab_n) %>%
      ungroup()
    cmpd_eic_i %>%
      inner_join(unlabeled_eic, by = join_by(rt, filename)) %>%
      group_by(filename, iso_name) %>%
      filter(n()>5) %>%
      arrange(filename, iso_name, rt) %>%
      summarise(cor_model=list(cor.test(int, unlab_int)), 
                cor_coef=cor_model[[1]]$estimate,
                cor_pval=cor_model[[1]]$p.value,
                overlap_frac=n()/unique(unlab_n), 
                area=trapz(rt, int, baseline="square"),
                .groups = "drop")
}, simplify = FALSE) %>% 
  bind_rows(.id = "compound_name") %>%
  group_by(compound_name) %>%
  mutate(fdr_pval=p.adjust(cor_pval, method="fdr")) %>%
  mutate(holm_pval=p.adjust(cor_pval, method="holm")) %>%
  mutate(compound_name=factor(compound_name, levels=unique(cmpd_eics$compound_name))) %>%
  arrange(compound_name) %>%
  ungroup()
```

```{r inspect raw peaks to decide on thresholds}
raw_areas %>%
  # filter(compound_name=="L-Arginine") %>%
  filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
  left_join(parametadata %>% distinct(filename, amendment, startime, depth, timepoint)) %>%
  arrange(timepoint) %>%
  filter(amendment=="Arg") %>%
  filter(startime=="Morn" & depth=="Surf") %>%
  distinct(filename, compound_name, amendment) %>%
  drop_na() %>% 
  # select(filename, overlap_frac, cor_coef) %>% print(n=Inf)
  pwalk(function(...){
    row_data <- data.frame(...)
    gp <- row_data %>% 
      left_join(cmpd_eics, by = join_by(compound_name, amendment, filename)) %>%
      group_by(iso_name) %>%
      ggplot() +
      geom_line(aes(x=rt, y=int, group = interaction(iso_name, filename), color=iso_name)) +
      facet_wrap(~iso_name, ncol=1, scales="free_y") +
      ggtitle(row_data$filename) +
      theme(legend.position="none")
    print(gp)
  }, .progress = TRUE)
raw_areas %>%
  # filter(compound_name=="L-Arginine") %>%
  filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
  left_join(parametadata %>% distinct(filename, amendment, startime, depth, shortname)) %>%
  filter(amendment=="Arg") %>%
  filter(startime=="Morn" & depth=="Surf") %>%
  drop_na() %>% 
  select(shortname, iso_name, overlap_frac, cor_coef, holm_pval, area) %>%
  print(n=Inf)
```

```{r define peak areas by filtering out intrusive ions}
deionized_areas <- raw_areas %>%
  filter(cor_pval<0.01) %>%
  select(filename, compound_name, iso_name, area)

raw_areas %>%
  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
  filter(compound_name=="L-Glutamine") %>%
  filter(amendment=="Arg") %>%
  filter(depth=="Deep") %>%
  filter(startime=="Morn") %>%
  filter(iso_name=="L-Glutamine, 13C5, 15N1") %>%
  select(filename, iso_name, cor_pval, area) %>%
  print(n=Inf)
```

```{r visualize final peak areas}
peak_areas %>%
  filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
  left_join(parametadata) %>%
  filter(samp_type=="Smp") %>%
  # filter(amendment%in%c("Arg", "GMP")) %>%
  ggplot() +
  geom_col(aes(x=tripl, y=area, fill=iso_name)) +
  facet_nested(depth+amendment~startime+timepoint, scales="free_y")
# cmpd_eics %>%
#   filter(compound_name=="L-Glutamic acid/beta-Glutamic acid") %>%
#   left_join(parametadata) %>%
#   filter(samp_type=="Smp") %>%
#   # filter(amendment%in%c("Arg", "GMP")) %>%
#   filter(timepoint=="T3") %>%
#   ggplot() +
#   geom_line(aes(x=rt, y=int, group=filename, color=amendment)) +
#   facet_wrap(~iso_name, ncol=2, scales="free_y")
```

### Fix problematic areas from IS contamination

```{r}
is_contam_shortnames <- c("Amm-Surf-Eve-T3-A", "Amm-Surf-Eve-T3-B", 
                          "Amm-Surf-Morn-T26-A", "Arg-Deep-Eve-T10-A", 
                          "Nit-Deep-Eve-T0-A")
problem_stans <- all_stans %>%
  filter(compound_type=="Internal Standard") %>%
  select(compound_name, polarity) %>%
  mutate(unlab_name=str_remove(compound_name, ", .*")) %>%
  filter(str_detect(compound_name, "15N|13C")) %>%
  mutate(n_C=str_extract(compound_name, "(?<=13C)\\d+")) %>%
  mutate(n_C=case_when(
    !str_detect(compound_name, "13C")~"0",
    str_detect(compound_name, "13C") & is.na(n_C)~"1",
    TRUE~n_C
  )) %>%
  mutate(n_N=str_extract(compound_name, "(?<=15N)\\d+")) %>%
  mutate(n_N=case_when(
    !str_detect(compound_name, "15N")~"0",
    str_detect(compound_name, "15N") & is.na(n_N)~"1",
    TRUE~n_N
  )) %>%
  mutate(contam_name=paste0(unlab_name, ", 13C", n_C, ", 15N", n_N)) %>%
  expand_grid(shortname=is_contam_shortnames) %>%
  left_join(parametadata %>% distinct(shortname, filename)) %>%
  select(iso_name=contam_name, filename)

decontam_areas <- deionized_areas %>%
  anti_join(problem_stans)
```

```{r write out}
write_csv(decontam_areas, "targeted/all_peak_areas_w_isos.csv")
```

### Some visualizations

```{r}
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv")
```

```{r Stacked barplots}
peak_areas %>%
  filter(compound_name%in%c("Theanine?")) %>%
  inner_join(parametadata %>% filter(samp_type=="Smp")) %>%
  ggplot() +
  geom_col(aes(x=tripl, y=area, fill=iso_name), position = position_fill(), 
           color="black", width=1) +
  facet_nested(amendment+depth~compound_name+startime+timepoint, scales="free_y")
```

```{r Logistic regression curves}
peak_areas %>%
  filter(compound_name=="L-Glutamine") %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name) %>%
  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
  filter(!is.na(startime)) %>%
  ggplot(aes(x=timepoint_num, y=labeled_frac)) +
  geom_point() +
  geom_smooth(method = "glm", formula = y~x, method.args = list(family = "quasibinomial")) +
  facet_nested(amendment+depth~startime)
```

```{r Logistic regression coefficients}
peak_areas %>%
  filter(compound_name=="L-Glutamine") %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment, depth, startime, timepoint_num)) %>%
  filter(samp_type=="Smp") %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name, -samp_type) %>%
  nest(data=c(filename, timepoint_num, labeled_frac)) %>%
  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
  unnest(coefs) %>%
  # filter(p.adjust(p.value, "fdr")<0.05) %>%
  select(-statistic, -p.value, -std.error) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
  mutate(half_max=-intercept/slope, scale=1/slope) %>%
  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24)))) %>%
  ggplot() +
  geom_point(aes(x=half_max, y=frac_at_T24, color=amendment, shape=depth, size=startime), alpha=0.5) +
  scale_size_manual(values=c(5, 3))
```

```{r Logreg attempting to propagate errors via delta correction}
library(msm)
peak_areas %>%
  filter(compound_name=="L-Glutamine") %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment, depth, startime, timepoint_num)) %>%
  filter(samp_type=="Smp") %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name, -samp_type) %>%
  nest(data=c(filename, timepoint_num, labeled_frac)) %>%
  mutate(log_fit=map(data, ~glm(.x$labeled_frac~.x$timepoint_num, family="quasibinomial")),
         coefs=map(log_fit, function(x)broom::tidy(x))) %>%
  unnest(coefs) %>%
  select(-statistic, -p.value, -std.error) %>%
  mutate(term=ifelse(term=="(Intercept)", "intercept", "slope")) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  mutate(half_max=-intercept/slope) %>%
  mutate(frac_at_T24 = 1/(1+exp(-(intercept+slope*24)))) %>%
  rowwise() %>%
  mutate(half_max_se = msm::deltamethod(~ -x1 / x2, coef(log_fit), vcov(log_fit))) %>%
  mutate(frac24_se = msm::deltamethod(~ 1 / (1 + exp(-(x1 + x2 * 24))), coef(log_fit), vcov(log_fit))) %>%
  ungroup() %>%
  ggplot(aes(x=half_max, y=frac_at_T24, color=amendment)) +
  geom_linerange(aes(ymin=frac_at_T24-2*frac24_se, ymax=frac_at_T24+2*frac24_se)) +
  geom_linerange(aes(xmin=half_max-2*half_max_se, xmax=half_max+2*half_max_se)) +
  geom_point(aes(shape=depth, size=startime), alpha=0.5) +
  scale_size_manual(values=c(5, 3))
```

```{r Logreg attempting error distributions via bootstrapping}
set.seed(123)
all_bootstraps <- peak_areas %>%
  filter(compound_name == "Glycine betaine") %>%
  mutate(labeled_frac = 1 - area / sum(area), .by = c("filename", "compound_name")) %>%
  select(compound_name, iso_name, filename, labeled_frac) %>%
  left_join(parametadata %>% distinct(filename, samp_type, amendment, depth, startime, timepoint_num)) %>%
  filter(samp_type == "Smp") %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name, -samp_type) %>%
  nest(data = c(filename, timepoint_num, labeled_frac)) %>%
  mutate(log_fit = map(data, ~glm(.x$labeled_frac ~ .x$timepoint_num, family = "quasibinomial"))) %>%
  mutate(bootstrap_results = map(data, function(data_i) {
    boot_samples <- map(1:100, function(i) {
      boot_data <- data_i[sample(nrow(data_i), replace = TRUE), ]
      boot_model <- glm(labeled_frac ~ timepoint_num, data = boot_data, family = "quasibinomial")
      half_max_boot <- -coef(boot_model)[1] / coef(boot_model)[2]
      frac24_boot <- 1 / (1 + exp(-(coef(boot_model)[1] + coef(boot_model)[2] * 24)))
      data.frame(half_max_boot=half_max_boot, frac24_boot=frac24_boot)
    }) %>%
      bind_rows()
  }))
# Plot raw data colored by amendment
all_bootstraps %>%
  unnest(bootstrap_results) %>%
  ggplot(aes(x = half_max_boot, y = frac24_boot, color = amendment)) +
  geom_point(alpha = 0.5) +
  scale_size_manual(values = c(5, 3))
# Plot summarized data as mean +/- SD
all_bootstraps %>%
  unnest(bootstrap_results) %>%
  summarise(half_max_mean=mean(half_max_boot), half_max_sd=sd(half_max_boot),
            frac24_mean=mean(frac24_boot), frac24_sd=sd(frac24_boot),
            .by = c(amendment, depth, startime)) %>%
  ggplot(aes(x=half_max_mean, y=frac24_mean, color=amendment, label=paste(depth, startime))) +
  geom_point() +
  geom_text(nudge_y = 0.02) +
  geom_linerange(aes(ymin=frac24_mean-2*frac24_sd, ymax=frac24_mean+2*frac24_sd)) +
  geom_linerange(aes(xmin=half_max_mean-2*half_max_sd, xmax=half_max_mean+2*half_max_sd))
# Plot raw data faceted by amendment and colored by exp factor
all_bootstraps %>%
  unnest(bootstrap_results) %>%
  ggplot(aes(x = half_max_boot, y = frac24_boot, color = interaction(depth, startime))) +
  geom_point(alpha = 0.5) +
  scale_size_manual(values = c(5, 3)) +
  facet_wrap(~amendment, scales="free")
```

```{r Plotting peak areas of each isotope as a heatmap instead}
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv")
peak_areas %>% 
  left_join(parametadata %>% filter(IS_type=="noIS")) %>%
  filter(samp_type=="Smp") %>%
  filter(compound_name=="Adenine") %>%
  left_join(iso_stans) %>%
  arrange(depth, startime, timepoint) %>%
  mutate(filename=fct_inorder(filename)) %>%
  mutate(area=area/sum(area), .by=filename) %>%
  mutate(n_N=factor(n_N, levels=5:0)) %>%
  mutate(n_C=factor(n_C, levels=10:0)) %>%
  filter(amendment%in%c("Amm", "Nit")) %>%
  # filter(samp_type=="Smp") %>%
  # filter(startime=="Morn") %>%
  # filter(n_N==1) %>%
  ggplot() +
  geom_tile(aes(x=timepoint, y=tripl, fill=area)) +
  scale_fill_viridis_c() +
  ggh4x::facet_nested(amendment+n_C+n_N~depth+startime, 
                      scales = "free", space = "free")
```

```{r Raw chromatogram to double check}
v <- open_dataset("tmzMLs/pqds") %>%
  filter(polarity=="pos") %>%
  filter(mz%between%pmppm(174.1365, 10)) %>%
  collect()

v %>%
  # filter(amendment=="Amm") %>% filter(int>1e6) %>% distinct(filename)
  filter(!filename%in%c(
    "210916_Smp_Amm-Surf-Eve-T3-A-noIS-pos.mzML",
    "210916_Smp_Amm-Surf-Eve-T3-B-noIS-pos.mzML",
    "210916_Smp_Amm-Surf-Morn-T26-A-noIS-pos.mzML"
  )) %>%
  filter(str_detect(filename, "Smp")) %>%
  left_join(parametadata) %>%
  qplotMS1data(color_col = "timepoint") +
  facet_wrap(~amendment, scales = "free_y")
```

```{r}
peak_areas %>%
  group_by(filename, compound_name) %>%
  mutate(area=area/sum(area)) %>%
  mutate(label_type=ifelse(str_detect(iso_name, ", 13C0, 15N0"), "unlabeled", "labeled")) %>%
  group_by(filename, compound_name, label_type) %>%
  summarise(area_unlabeled=sum(area)) %>%
  filter(label_type=="unlabeled") %>%
  left_join(parametadata) %>%
  filter(startime=="Morn") %>%
  filter(depth=="Deep") %>%
  # filter(compound_name=="L-Glutamine") %>%
  group_by(compound_name, amendment, timepoint) %>%
  summarise(labeled_above_thresh=median(area_unlabeled)<0.5) %>% #Less than X peak area is in unlabeled fraction
  group_by(amendment, timepoint) %>%
  summarise(n_labeled=sum(labeled_above_thresh), n_total=n()) %>%
  select(amendment, timepoint, n_labeled, total=n_total) %>%
  ungroup() %>%
  mutate(total=median(total), .by = amendment) %>%
  pivot_wider(names_from = timepoint, values_from = n_labeled)
```

## PheTau things (not currently used for anything)

Just confirming that all the injections went off successfully

```{r IS areas}
IS_eics <- open_dataset("tmzMLs/pqds") %>%
      filter(mz%between%c(128.0308598, 128.0321402) |
               mz%between%c(130.0464498, 130.0477502) |
               mz%between%c(174.1356293, 174.1373707)) %>%
      collect()

IS_bounds <- tribble(
  ~amendment, ~polarity, ~compound_name, ~mz, ~rtstart, ~rtend, 
  "Amm", "neg", "Taurine, 2H4", 128.0315, 9.8, 11.5,
  "Arg", "neg", "Taurine, 2H4", 128.0315, 8.5, 11,
  "GMP", "neg", "Taurine, 2H4", 128.0315, 9.5, 11.5,
  "Nit", "neg", "Taurine, 2H4", 128.0315, 10, 10.5,
  "Amm", "pos", "Taurine, 2H4", 130.0471, 9.8, 11.5,
  "Arg", "pos", "Taurine, 2H4", 130.0471, 8.5, 11,
  "GMP", "pos", "Taurine, 2H4", 130.0471, 9.5, 11.5,
  "Nit", "pos", "Taurine, 2H4", 130.0471, 10, 10.5,
  "Amm", "pos", "L-Phenylalanine, 2H8", 174.1365, 4, 6.5,
  "Arg", "pos", "L-Phenylalanine, 2H8", 174.1365, 3.6, 6,
  "GMP", "pos", "L-Phenylalanine, 2H8", 174.1365, 4, 7.2,
  "Nit", "pos", "L-Phenylalanine, 2H8", 174.1365, 3.5, 5
)

IS_areas <- IS_bounds %>%
  rowwise() %>%
  mutate(mzmin = pmppm(mz, 5)[1], mzmax = pmppm(mz, 5)[2]) %>%
  ungroup() %>%
  select(-mz) %>%
  left_join(IS_eics, by=join_by(
      amendment, polarity, dplyr::between(y$rt, x$rtstart, x$rtend), 
      dplyr::between(y$mz, x$mzmin, x$mzmax)
    )) %>%
  distinct() %>%
  # qplotMS1data(color_col = "amendment") + facet_wrap(~compound_name+polarity, ncol = 1, scales = "free")
  group_by(compound_name, amendment, filename, rt) %>%
  slice_max(int) %>%
  group_by(amendment, filename, compound_name, polarity) %>%
  arrange(rt) %>%
  summarise(area=trapz(rt, int), .groups="drop")

bad_IS_files <- c(
  "210916_Smp_Amm-Deep-Eve-T1-C-noIS",
  "220729_Smp_Arg-Deep-Eve-T10-A-noIS",
  "220715_Smp_GMP-Deep-Eve-T26-B-noIS",
  "220715_Smp_GMP-Surf-Eve-T3-C-noIS",
  "220729_Smp_Arg-Surf-Morn-T73-C-noIS"
) %>%
  rep(each=2) %>%
  paste0(c("-pos.mzML", "-neg.mzML"))

fixed_IS_areas <- IS_areas %>%
  left_join(parametadata) %>%
  filter(samp_type=="Smp") %>%
  arrange(compound_name, polarity, amendment, timestamp) %>%
  group_by(compound_name, polarity, amendment) %>%
  mutate(lag_area=lag(area), lead_area=lead(area)) %>%
  mutate(area=ifelse(filename%in%bad_IS_files, (lag_area+lead_area)/2, area)) %>%
  select(amendment, filename, compound_name, polarity, area)
IS_areas <- rows_update(IS_areas, fixed_IS_areas, by = c("amendment", "filename", "compound_name", "polarity"))

IS_areas %>%
  select(amendment, filename, compound_name, polarity, area) %>%
  filter(amendment!="Nit") %>%
  write_csv("targeted/PheTau_peak_areas.csv")
```

```{r vis IS areas}
IS_areas <- read_csv("targeted/PheTau_peak_areas.csv")
IS_areas %>%
  left_join(parametadata) %>%
  arrange(timestamp) %>%
  mutate(filename=str_remove(filename, "-pos|-neg")) %>%
  mutate(filename=fct_inorder(filename)) %>%
  filter(samp_type%in%c("Poo", "Smp")) %>%
  ggplot() +
  geom_col(aes(x=filename, y=area, fill=amendment), width=1) +
  facet_grid(compound_name+polarity~samp_type, scales = "free", space = "free_x")

# Check for missing samples
# parametadata %>%
#   filter(IS_type=="noIS") %>%
#   anti_join(distinct(IS_areas, filename)) %>%
#   distinct(filename) %>%
#   print(n=Inf)

# Test correlation between IS
# IS_areas %>%
#   filter(amendment!="Nit") %>%
#   filter(str_detect(filename, "Poo")) %>%
#   mutate(filename=str_remove(filename, "-pos|-neg")) %>%
#   pivot_wider(names_from=c(compound_name, polarity), values_from = area) %>%
#   # select(3:5) %>% cor() %>% round(3) %>% `diag<-`("")
#   ggplot() +
#   geom_point(aes(x=`Taurine, 2H4_neg`, y=`L-Phenylalanine, 2H8_pos`, color=amendment))
```

```{r}
IS_areas <- read_csv("targeted/PheTau_peak_areas.csv")
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv")

pooled_metadata <- parametadata %>%
  filter(samp_type=="Poo") %>%
  filter(IS_type=="noIS") %>%
  filter(amendment!="Nit") %>%
  select(filename)
pooled_metadata %>%
  left_join(IS_areas)
pooled_metadata %>%
  left_join(peak_areas)
```

## Quant (requires starting over for the wIS samples)

### Get initial peak areas (for both IS and all compounds)

```{r plot chroms to estimate peak bounds, eval=FALSE}
is_chroms <- read_csv("metadata/all_stans.csv") %>%
  filter(compound_type=="Internal Standard") %>%
  filter(date_added<200101) %>%
  select(compound_name, polarity, mz) %>%
  mutate(is_name=compound_name) %>%
  mutate(compound_name=str_remove(compound_name, ",.*")) %>%
  rename(is_mz=mz) %>%
  left_join(readxl::read_xlsx("targeted/peak_bounds.xlsx")) %>%
  filter(compound_name=="L-Cysteic acid, 2H3")

  drop_na("amendment") %>%
  select(compound_name=is_name, mz=is_mz, amendment, polarity, rtstart, rtend) %>%
  rowwise() %>%
  mutate(mzmin=pmppm(mz, 10)[1]) %>%
  mutate(mzmax=pmppm(mz, 10)[2]) %>%
  ungroup() %>%
  # slice(1) -> row_data
  pmap(function(...){
    row_data <- list(...)
    open_dataset("tmzMLs/pqds_wIS") %>%
      filter(amendment==row_data$amendment) %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(row_data$mzmin, row_data$mzmax)) %>%
      filter(rt%between%c(row_data$rtstart, row_data$rtend)) %>%
      collect() %>%
      filter(str_detect(filename, "T0")) %>%
      slice_max(int, by=c(filename, rt)) %>%
      mutate(compound_name=row_data$compound_name,
             polarity=row_data$polarity) %>%
      select(compound_name, amendment, filename, rt, int, polarity)
  }, .progress = TRUE) %>%
  bind_rows()
is_chroms %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  # filter(compound_name==unique(compound_name)[3]) %>%
  qplotMS1data(color_col="amendment") +
  facet_wrap(~compound_name, scales = "free")
plotly::ggplotly()

read_csv("metadata/all_stans.csv") %>%
  filter(compound_type=="Internal Standard") %>%
  filter(date_added<200101) %>%
  select(compound_name, polarity, mz) %>%
  mutate(is_name=compound_name) %>%
  mutate(compound_name=str_remove(compound_name, ",.*")) %>%
  rename(is_mz=mz) %>%
  left_join(readxl::read_xlsx("targeted/peak_bounds.xlsx")) %>%
  drop_na("amendment") %>%
  select(compound_name=is_name, mz=is_mz, amendment, polarity, rtstart, rtend) %>%
  write_csv("targeted/IS_bounds.csv")
```

```{r integrate IS areas, eval=FALSE}
is_chroms <- readxl::read_xlsx("targeted/IS_bounds.xlsx") %>%
  rowwise() %>%
  mutate(mzmin=pmppm(mz, 10)[1]) %>%
  mutate(mzmax=pmppm(mz, 10)[2]) %>%
  ungroup() %>%
  # slice(1) -> row_data
  pmap(function(...){
    row_data <- list(...)
    open_dataset("tmzMLs/pqds_wIS") %>%
      filter(amendment==row_data$amendment) %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(row_data$mzmin, row_data$mzmax)) %>%
      filter(rt%between%c(row_data$rtstart, row_data$rtend)) %>%
      collect() %>%
      # distinct(filename)
      slice_max(int, by=c(filename, rt)) %>%
      mutate(compound_name=row_data$compound_name, 
             polarity=row_data$polarity) %>%
      select(compound_name, amendment, filename, rt, int, polarity)
  }, .progress = TRUE) %>%
  bind_rows()
is_chroms %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  # filter(compound_name==unique(compound_name)[3]) %>%
  qplotMS1data(color_col="amendment") +
  facet_wrap(~compound_name, scales = "free")

is_areas <- is_chroms %>%
  arrange(rt) %>%
  summarise(IS_area=trapz(rt, int), .by=c(amendment, filename, compound_name, polarity)) %>%
  left_join(parametadata %>% distinct(filename, timestamp, samp_type, polarity)) %>%
  complete(filename, compound_name) %>%
  arrange(filename) %>%
  group_by(filename) %>%
  fill(timestamp, amendment, samp_type, .direction="downup") %>%
  ungroup() %>%
  arrange(compound_name, timestamp) %>%
  # filter(filename=="210813_Smp_Nit-Deep-Eve-T0-A-wIS.mzML")
  mutate(IS_area=case_when(
    filename=="210813_Smp_Nit-Deep-Eve-T0-A-wIS.mzML"~(lag(IS_area)+lead(IS_area))/2,
    TRUE~IS_area
  )) %>%
  select(amendment, filename, polarity, IS_name=compound_name, IS_area)
write_csv(is_areas, "targeted/IS_areas.csv")
```

```{r integrate nonIS unlabeled areas from wIS files}
wIS_bounds <- readxl::read_xlsx("targeted/peak_bounds.xlsx") %>%
  filter(rtstart>0) %>%
  filter(is.na(notes) | notes!="Remove") %>%
  filter(compound_name!="Homarine") %>% # Can't separate from Adenine IS sufficiently
  select(amendment, polarity, compound_name, mz, rtstart, rtend)

wIS_eics <- wIS_bounds %>%
  nest(amend_specs=-c(compound_name, polarity)) %>%
  # head() %>%
  # slice(1)  -> row_data
  # filter(compound_name=="Theanine?") -> row_data
  pmap(function(...){
    row_data <- list(...)
    # print(row_data$compound_name)
    amend_specs <- row_data$amend_specs %>%
      rowwise() %>%
      mutate(mzmin=pmppm(mz, 10)[1]) %>%
      mutate(mzmax=pmppm(mz, 10)[2]) %>%
      ungroup()
    mzmin <- min(amend_specs$mz)-0.5
    mzmax <- max(amend_specs$mz)+0.5
    rtmin <- min(amend_specs$rtstart)
    rtmax <- max(amend_specs$rtend)
    all_eic <- open_dataset("tmzMLs/pqds_wIS") %>%
      filter(polarity==row_data$polarity) %>%
      filter(mz%between%c(mzmin, mzmax)) %>%
      filter(rt%between%c(rtmin, rtmax)) %>%
      collect()
    cmpd_eic <- left_join(amend_specs, all_eic, by=join_by(
      amendment, dplyr::between(y$rt, x$rtstart, x$rtend), 
      dplyr::between(y$mz, x$mzmin, x$mzmax)
    )) %>%
      distinct() %>%
      group_by(amendment, filename, rt) %>%
      # nest() %>% ungroup() %>% slice(1) %>% unnest(data)
      slice_max(int) %>%
      ungroup() %>%
      mutate(compound_name=row_data$compound_name) %>%
      select(amendment, compound_name, rt, mz=mz.y, int, filename, polarity)
    cmpd_eic
  }, .progress = TRUE) %>%
  bind_rows() %>%
  ungroup()
saveRDS(wIS_eics, "targeted/wIS_eics.rds")

wIS_eics <- readRDS("targeted/wIS_eics.rds")
wIS_areas <- wIS_eics %>%
  slice_max(int, by=c(filename, compound_name, rt)) %>%
  arrange(rt) %>%
  summarise(area=trapz(rt, int, baseline = "square"), 
            .by = c(amendment, filename, polarity, compound_name)) %>%
  mutate(compound_name=factor(compound_name, levels=unique(wIS_eics$compound_name))) %>%
  arrange(compound_name, filename)
wIS_areas %>%
  write_csv("targeted/wIS_areas.csv")
```

### Refine IS peak areas due to contamination

```{r Replace wonky values with average of previous and next}
is_areas <- read_csv("targeted/IS_areas.csv")

IS_mix_one <- c("Thymine, 2H4", "Uracil, 15N2, 2H2", "Adenine, 15N2",
                "Cytosine, 13C2, 15N3", "Glycine betaine, 13C5, 15N",
                "Guanine, 13C, 15N2", "Homarine, 2H3")

fixed_IS <- parametadata %>%
  filter(samp_type%in%c("Smp", "Poo", "Blk")) %>%
  filter(IS_type=="wIS") %>%
  distinct(filename, shortname, timestamp) %>%
  left_join(is_areas %>% select(filename, amendment, IS_name, IS_area)) %>%
  arrange(IS_name, timestamp) %>%
  mutate(filename=fct_inorder(filename)) %>%
  # filter(shortname=="Nit-Deep-Eve-T1-A") %>%
  mutate(IS_area=case_when(
    shortname=="Nit-Deep-Eve-T1-A" & IS_name%in%IS_mix_one ~ 
      (lag(IS_area) + lead(IS_area))/2,
    shortname=="Amm-Surf-Eve-T3-A" ~ (lag(IS_area) + lead(IS_area))/2,
    shortname=="Arg-Deep-Eve-T10-A" ~ (lag(IS_area) + lead(IS_area))/2,
    shortname=="GMP-Surf-Eve-T3-C" ~ (lag(IS_area) + lead(IS_area, 2))/2,
    shortname=="GMP-Surf-Eve-T3-C" & IS_name=="L-Histidine, 15N" ~ 
      (lag(IS_area) + lead(IS_area, 2))/2,
    shortname=="GMP-Surf-Morn-T26-B" & IS_name=="L-Histidine, 15N" ~ 
      (lag(IS_area, 2) + lead(IS_area))/2,
    shortname=="Arg-Surf-Eve-T10-B" & IS_name=="Sulfoacetic acid, 13C2" ~ 
      (lag(IS_area) + lead(IS_area, 2))/2,
    shortname=="Arg-Deep-Eve-T3-A" & IS_name=="Sulfoacetic acid, 13C2" ~ 
      (lag(IS_area) + lead(IS_area, 2))/2,
    TRUE ~ IS_area
  ))
fixed_IS %>%
  select(filename, amendment, IS_name, IS_area) %>%
  write_csv("targeted/IS_areas.csv")
```

```{r double-checking that IS fixes worked via plots and PCA}
fixed_IS %>%
  # filter(str_detect(IS_name, "Homarine")) %>%
  left_join(parametadata %>% distinct(filename, timestamp, amendment)) %>%
  # filter(amendment=="Amm") %>%
  arrange(timestamp) %>%
  mutate(filename=fct_inorder(filename)) %>%
  ggplot() +
  geom_col(aes(x=filename, y=IS_area, fill=amendment), width=1) +
  facet_wrap(~IS_name, ncol=3, scales="free_y")

wide_IS <- fixed_IS %>%
  select(shortname, IS_name, IS_area) %>%
  # filter(!str_detect(shortname, "PooGMP")) %>%
  pivot_wider(names_from=IS_name, values_from=IS_area) %>%
  column_to_rownames("shortname")

# wide_IS %>% write.csv("~/../Desktop/fixed_IS.csv")

wide_IS %>%
  data.matrix() %>%
  prcomp() %>%
  pluck("x") %>%
  as.data.frame() %>%
  rownames_to_column("shortname") %>%
  left_join(parametadata %>% filter(IS_type=="wIS")) %>%
  ggplot() +
  geom_text(aes(x=PC1, y=PC2, label=filename, color=amendment))
wide_IS %>%
  data.matrix() %>%
  t() %>%
  prcomp() %>%
  pluck("x") %>%
  as.data.frame() %>%
  rownames_to_column("IS_name") %>%
  ggplot() +
  geom_text(aes(x=PC1, y=PC2, label=IS_name))

wide_IS %>%
  data.matrix() %>%
  t() %>%
  prcomp() %>%
  pluck("rotation") %>%
  as.data.frame() %>%
  rownames_to_column("shortname") %>%
  select(shortname, PC1:PC3) %>%
  pivot_longer(PC1:PC3) %>%
  left_join(parametadata %>% filter(IS_type=="wIS")) %>%
  ggplot() +
  geom_col(aes(x=shortname, y=value, fill=amendment), width=1) +
  facet_wrap(~name,ncol=1)
```

### Choose BMIS

```{r identify the BMIS for each compound}
IS_areas <- read_csv("targeted/IS_areas.csv")

chosen_IS <- IS_areas %>%
  inner_join(parametadata %>% distinct(filename, amendment, samp_type)) %>%
  filter(samp_type=="Poo") %>%
  select(filename, IS_name, amendment, IS_area) %>%
  left_join(wIS_areas, relationship = "many-to-many") %>%
  arrange(compound_name, filename) %>%
  group_by(amendment, compound_name, IS_name) %>%
  mutate(norm_area=area/IS_area*mean(IS_area)) %>%
  summarise(cv=sd(norm_area)/mean(norm_area), .groups = "drop") %>%
  slice_min(cv, by = c(compound_name, amendment)) %>%
  # print(n=Inf)
  # with(table(IS_name, compound_name))
  select(amendment, compound_name, IS_name)
count(chosen_IS, amendment, IS_name) %>%
  pivot_wider(names_from=amendment, values_from=n) %>%
  print(n=Inf)

IS_with_unlabeled <- IS_areas %>%
  distinct(IS_name, amendment) %>%
  mutate(compound_name=str_remove(IS_name, ", .*"))
chosen_IS %>%
  filter(compound_name%in%IS_with_unlabeled$compound_name) %>%
  count(compound_name, IS_name) %>%
  print(n=Inf)

chosen_IS <- chosen_IS %>%
  rows_update(IS_with_unlabeled, by = c("compound_name", "amendment"), unmatched = "ignore")
write_csv(chosen_IS, "targeted/chosen_IS.csv")
```

### Calculate RFs

```{r create response factor dataframe from standards}
wIS_areas <- read_csv("targeted/wIS_areas.csv")
stan_patched <- all_stans %>%
  filter(!compound_name=="beta-Glutamic acid") %>%
  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
  filter(!compound_name=="L-Isoleucine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
  filter(!compound_name=="L-Homoserine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
  filter(!compound_name=="Allopurinol") %>%
  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))

rf_df <- wIS_areas %>%
  left_join(parametadata %>% select(filename, samp_type, polarity)) %>%
  filter(samp_type=="Std") %>%
  left_join(stan_patched %>% distinct(compound_name, mix, conc_um)) %>%
  filter(str_detect(filename, paste0(mix, "InMatrix")) | 
           str_detect(filename, "H2OInMatrix")) %>%
  select(compound_name, conc_um, amendment, filename, area) %>%
  mutate(filename=str_extract(filename, "Mix|H2O")) %>%
  pivot_wider(names_from = filename, values_from=area, values_fn = mean) %>%
  mutate(area_diff=Mix-H2O) %>%
  # filter(rf<0)
  filter(area_diff>0) %>%
  mutate(rf=area_diff/conc_um) %>%
  select(compound_name, amendment, rf)

write_csv(rf_df, "targeted/rf_df.csv")
rf_df %>%
  arrange(rf) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  ggplot() +
  geom_boxplot(aes(x=rf, y=compound_name)) +
  scale_x_log10()
```

### Finalize concentrations and drop low-quality compounds

```{r quantify}
wIS_areas <- read_csv("targeted/wIS_areas.csv")
chosen_IS <- read_csv("targeted/chosen_IS.csv")
IS_areas <- read_csv("targeted/IS_areas.csv")
rf_df <- read_csv("targeted/rf_df.csv")

env_concs <- wIS_areas %>%
  filter(!str_detect(compound_name, "\\?$")) %>%
  filter(!compound_name%in%c("Urea", "L-Valine", "Oxalic acid", "Malic acid", 
                             "Acetylcholine", "Homarine")) %>%
  inner_join(parametadata %>% filter(samp_type=="Smp") %>% distinct(filename)) %>%
  left_join(chosen_IS) %>%
  left_join(IS_areas) %>%
  mutate(norm_area=area/IS_area*mean(IS_area), .by = c(amendment, compound_name)) %>%
  left_join(rf_df) %>%
  mutate(conc_um_in_vial=norm_area/rf) %>%
  select(compound_name, filename, conc_um_in_vial) %>%
  mutate(conc_nm_in_env=conc_um_in_vial*400/2e6*1000)
write_csv(env_concs, "targeted/final_env_concs.csv")
```

```{r quant plots for confirmation}
env_concs %>%
  filter(compound_name%in%c("Gonyol", "L-Glutamine", "Guanosine",
                            "L-Glutamic acid/beta-Glutamic acid")) %>%
  left_join(parametadata, relationship = "many-to-many") %>%
  ggplot() +
  geom_hline(yintercept = 0) +
  geom_boxplot(aes(x=amendment, y=conc_nm_in_env, color=depth)) +
  facet_wrap(~compound_name, scales="free_y")

env_concs %>%
  filter(conc_nm_in_env==0) %>%
  left_join(parametadata, relationship = "many-to-many") %>%
  count(compound_name, amendment) %>%
  complete(compound_name, amendment, fill = list(n=0)) %>%
  pivot_wider(names_from = amendment, values_from=n)

env_concs %>%
  mutate(avg_conc=mean(conc_nm_in_env, na.rm=TRUE), .by=compound_name) %>%
  arrange(avg_conc) %>%
  mutate(compound_name=fct_inorder(compound_name)) %>%
  left_join(parametadata, relationship = "many-to-many") %>%
  filter(conc_nm_in_env>0) %>%
  ggplot() +
  geom_boxplot(aes(x=conc_nm_in_env, y=compound_name, color=depth)) +
  scale_x_log10(breaks=c(0.000001, 0.001, 1), labels=c("1 fM", "1 pM", "1 nM"))
```